---
title: "TASK 5: Exploratory Data Analysis - Sports"
author: "Yogesh S"
date: "2023-06-18"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
fig_width: 100
fig_height: 8
---

```{r,echo=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8)
```

<br>

##### **OBJECTIVE 1**: Perform Exploratory Data Analysis on dataset "Indian Premier League".

##### **OBJECTIVE 2**: Find out the most Successful teams.

##### **OBJECTIVE 3**: Identify players contributing to the team's success.

##### **OBJECTIVE 4**: Suggest Teams and players for companies to endorse their products.

<br>

### Let us perform Exploratory Data Analysis on the Dataset (Objective 1)

## 

#### **Step 1: Load the new Datasets.**

```{r}
setwd("C:/R/R Sample Datasets")
library(readr)
library(ggplot2)
myData<- read_csv("deliveries.csv")
myData2<- read_csv("matches.csv")
```

## 

<br>

#### **Step 2: View Datasets.**

```{r}
View(myData)
head(myData)
View(myData2)
head(myData2)
```

<br>

#### **Step 3: Check for missing values in the datasets using colSums().**

<br>

##### Let us Check for missing values in "MATCHES" dataset.

```{r}
colSums(is.na(myData))
```

##### The last 3 columns have missing values\>90%

## 

##### ***METHOD1: Removal of largely inconsistent data***

##### Let us remove these inconsistent data columns using the following code:

<br>

```{r}
subs1<-subset(myData, select = -c(player_dismissed,dismissal_kind,fielder))
colSums(is.na(subs1))
```

##### The dataset is now clear from any missing data.

<br>

##### Now let us Check for missing values in "MATCHES" dataset.

```{r}
colSums(is.na(myData2))
```

###### Remove the columns that have (missing values at random.)

```{r}
subs2<-subset(myData2, select = -c(umpire3))
colSums(is.na(subs2))
```

<br>

##### ***METHOD2: Replace value with other column value if NA values are found.***

<br>

##### Select subset containing NA values in the city column and select a replacement column.

<br>

```{r}
library(dplyr)
test <- subs2 %>% filter(is.na(city),)
select(test,city,venue)
subs2$city <- ifelse(is.na(subs2$city), subs2$venue, subs2$city)
colSums(is.na(subs2))
subs2<-na.omit(subs2)
colSums(is.na(subs2))
```

<br>

#### **Step 4: Summarize and understand the dataset better before analysis and visualization.**

```{r}
str(subs1)
str(subs2)
summary(subs1)
summary(subs2)
```

<br>

#### **Step 5: Analyze the data using Descriptive Statistics.**

```{r}
subset2<-filter(subs1,batsman=='MS Dhoni')
subset4<-filter(subs2,winner=='Chennai Super Kings')
mean(subset2$total_runs)
mean(subset4$win_by_runs)
var(subset4$win_by_wickets)
median(subset2$total_runs)
```

<br>

#### **Step 6: VISUALIZATION: We shall start Visualizing the Data when we solve the objectives of the task.**

<br>

### OBJECTIVE 2: Find the most successful teams

```{r}
q<-ggplot(subs2, aes(x=winner)) + geom_bar()
q + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

#### *According to the visualization, the most successful teams in IPL rank in the order:*

##### 1. Mumbai Indians

##### 2. Chennai Super Kings

##### 3. Kolkata Knight Riders

##### 4. Royal Challengers Bangalore

<br>

### OBJECTIVE 3: Find the players contributing to the success of each team

##### ***Each team is colour coded according to their 'JERSEY'***

<br>

### ***Mumbai Indians***

```{r}
MI_POM<-filter(subs2,winner=='Mumbai Indians')
MiMoM<-ggplot(MI_POM, aes(x=player_of_match)) + geom_bar(fill="blue")
MiMoM + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

<br>

##### **Player contributing most to the team's success: RG Sharma**

<br>

### ***Chennai Super Kings***

```{r}
CSK_POM<-filter(subs2,winner=='Chennai Super Kings')
CskMoM<-ggplot(CSK_POM, aes(x=player_of_match)) + geom_bar(fill="yellow")
CskMoM + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

<br>

##### **Player contributing most to the team's success: MS Dhoni**

<br>

### ***Kolkata Knight Riders***

```{r}
KKR_POM<-filter(subs2,winner=='Kolkata Knight Riders')
KkrMoM<-ggplot(KKR_POM, aes(x=player_of_match)) + geom_bar(fill="purple")
KkrMoM + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

<br>

##### **Player contributing most to the team's success: AD Russell**

<br>

### ***Deccan Chargers***

```{r}
DEC_POM<-filter(subs2,winner=='Deccan Chargers')
DecMoM<-ggplot(DEC_POM, aes(x=player_of_match)) + geom_bar(fill="navy")
DecMoM + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

<br>

##### **Player contributing most to the team's success: AC Gilchrist**

<br>

### ***Delhi Capitals***

```{r}
DC_POM<-filter(subs2,winner=='Delhi Capitals')
DcMoM<-ggplot(DC_POM, aes(x=player_of_match)) + geom_bar(fill="royalblue")
DcMoM + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

<br>

##### **Player contributing most to the team's success: RR Pant & S Dhawan**

<br>

### ***Delhi Daredevils***

```{r}
DD_POM<-filter(subs2,winner=='Delhi Daredevils')
DdMoM<-ggplot(DD_POM, aes(x=player_of_match)) + geom_bar(fill="red")
DdMoM + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

<br>

##### **Player contributing most to the team's success: V Sehwag**

<br>

### ***Gujarat Lions***

```{r}
GL_POM<-filter(subs2,winner=='Gujarat Lions')
GlMoM<-ggplot(GL_POM, aes(x=player_of_match)) + geom_bar(fill="orange")
GlMoM + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

<br>

##### **Player contributing most to the team's success: AJ Finch & DR Smith**

<br>

### ***Kings XI Punjab***

```{r}
KXIP_POM<-filter(subs2,winner=='Kings XI Punjab')
KxipMoM<-ggplot(KXIP_POM, aes(x=player_of_match)) + geom_bar(fill="brown2")
KxipMoM + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

<br>

##### **Player contributing most to the team's success: SE Marsh**

<br>

### ***Kochi Tuskers Kerala***

```{r}
KTK_POM<-filter(subs2,winner=='Kochi Tuskers Kerala')
KtkMoM<-ggplot(KTK_POM, aes(x=player_of_match)) + geom_bar(fill="black")
KtkMoM + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

<br>

##### **Player contributing most to the team's success: BB McCullum & BJ Hodge**

<br>

### ***Pune Warriors***

```{r}
PW_POM<-filter(subs2,winner=='Pune Warriors')
PwMoM<-ggplot(PW_POM, aes(x=player_of_match)) + geom_bar(fill="violet")
PwMoM + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

<br>

##### **Player contributing most to the team's success: SPD Smith**

<br>

### ***Rajasthan Royals***

```{r}
RR_POM<-filter(subs2,winner=='Rajasthan Royals')
RrMoM<-ggplot(RR_POM, aes(x=player_of_match)) + geom_bar(fill="cyan")
RrMoM + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

<br>

##### **Player contributing most to the team's success: AM Rahane**

<br>

### ***Rising Pune SupergiantS***

```{r}
RPS_POM<-filter(subs2,winner=='Rising Pune Supergiant')
RpsMoM<-ggplot(RPS_POM, aes(x=player_of_match)) + geom_bar(fill="pink")
RpsMoM + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

<br>

##### **Player contributing most to the team's success: BA Stokes**

<br>

### ***Royal Challengers Bangalore***

```{r}
RCB_POM<-filter(subs2,winner=='Royal Challengers Bangalore')
RcbMoM<-ggplot(RCB_POM, aes(x=player_of_match)) + geom_bar(fill="darkred")
RcbMoM + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

<br>

##### **Player contributing most to the team's success: AB de Villiers**

<br>

### ***Sunrisers Hyderabad***

```{r}
SH_POM<-filter(subs2,winner=='Sunrisers Hyderabad')
ShMoM<-ggplot(SH_POM, aes(x=player_of_match)) + geom_bar(fill="darkorange")
ShMoM + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

<br>

##### **Player contributing most to the team's success: DA Warner**

<br>

### OBJECTIVE 4: Suggest players or teams for a company to endorse for its products

#### Let us rerun our result from OBJECTIVE 2 to select successful teams to suggest.

```{r}
q + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

<br>

#### The successful teams are:

##### **1.Mumbai Indians**

##### **2.Chennai Super Kings**

##### **3.Kolkata Knight Riders**

##### **4.Royal Challengers Bangalore**

<br>

#### Let us analyze the popularity of players in the above teams by measuring the total runs scored by them

#### **Higher Performance of the player = Greater Popularity**

```{r}
library(dplyr)
```

<br>

### ***Mumbai Indians***

```{r}
MI<-filter(subs1,batting_team=='Mumbai Indians')
MI_tbl <- MI %>% group_by(batsman) %>% 
summarise(total_count=n(),
.groups = 'drop')
MI_tbl
MI2 <- MI_tbl %>% as.data.frame()
MI2 
MIplot<- ggplot(MI2, aes(x=batsman, y=total_count))+geom_bar(width = 1, stat = "identity",fill="blue")
MIplot + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

<br>

##### **Players with most popularity for ads are:**

##### 1. RG Sharma

##### 2. SR Tendulkar

##### 3. KA Pollard

<br>

### ***Chennai Super Kings***

```{r}
CSK<-filter(subs1,batting_team=='Chennai Super Kings')
CSK_tbl <- CSK %>% group_by(batsman) %>% 
  summarise(total_count=n(),
            .groups = 'drop')
CSK_tbl
# Convert tibble to df
CSK2 <- CSK_tbl %>% as.data.frame()
CSK2 
CSKplot<- ggplot(CSK2, aes(x=batsman, y=total_count))+geom_bar(width = 1, stat = "identity",fill="yellow")
CSKplot + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

<br>

##### **Players with most popularity for ads are:**

##### 1. SK Raina

##### 2. MS Dhoni

##### 3. MEK Hussey

<br>

### ***Kolkata Knight Riders***

```{r}
KKR<-filter(subs1,batting_team=='Kolkata Knight Riders')
KKR_tbl <- KKR %>% group_by(batsman) %>% 
  summarise(total_count=n(),
            .groups = 'drop')
KKR_tbl
# Convert tibble to df
KKR2 <- KKR_tbl %>% as.data.frame()
KKR2 
KKRplot<- ggplot(KKR2, aes(x=batsman, y=total_count))+geom_bar(width = 1, stat = "identity",fill="purple")
KKRplot + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

<br>

##### **Players with most popularity for ads are:**

##### 1. G Gambhir

##### 2. RV Uthappa

##### 3. YK Pathan

<br>

### ***Royal Challengers Bangalore***

```{r}
RCB<-filter(subs1,batting_team=='Royal Challengers Bangalore')
RCB_tbl <- RCB %>% group_by(batsman) %>% 
  summarise(total_count=n(),
            .groups = 'drop')
RCB_tbl
# Convert tibble to df
RCB2 <- RCB_tbl %>% as.data.frame()
RCB2 
RCBplot<- ggplot(RCB2, aes(x=batsman, y=total_count))+geom_bar(width = 1, stat = "identity",fill="darkred")
RCBplot + theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

<br>

##### **Players with most popularity for ads are:**

##### 1. V Kohli

##### 2. AB de Villiers

##### 3. CH Gayle

<br>

### RESULT:

<br>

#### Thus, we have successfully completed an extensive "Exploratory Data Analysis" using the dataset "Indian Premier League"

<br>

#### We have also analyzed successful teams and suggested players and teams that can be used by companies to endorse their products.
